### 2016-11-13 presentation.Rmd
added presentation

---
title: "Facial Recognition"
author: 
  - Stephanie Kobakian
  - Mitchell O'Hara-Wild
date: "26 September 2016"
output: 
  revealjs::revealjs_presentation:
    theme: black
    highlight: pygments
    transition: slide
    center: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Winter research project {data-background="img/intro_bg.png"}

## Project goals

- Find the best facial recognition software for tennis matches
- Build manually classified dataset of faces
- Compare software performance under varied conditions
- Determine ideal camera feeds for facial recognition

## Challenges

- Access to usable images and videos
- Availability of software
- Using the software
- Time constraints

# Applications of facial detection software

## Camera auto-focus
![Camera auto-focus](img/Face_detection.jpg)

## Facebook's image tagging
![Facebook's tagging feature for quick classification of friends](img/FacebookTag.png)

## Snapchat filters
![snap](img/snapchat.jpg)

## Surveillance and security
![security](img/security.png)

## Google's Mobile Vision

![Google Mobile Vision](img/googleMobile.png)

# Applying facial recognition to tennis

## Emotion in tennis

Serena Williams

> "Tennis is mostly mental. You win or lose the match before you even go out there."

![Serena GIF](img/serenaWilliams.gif)

## Long-term goals

- Better understand emotion's effect on player performance
- Automatic collection of player emotion data from video

# Manual classification

## Dataset

- 6406 Australian Open images (2.8GB)
- 800x450px size frames from 105 match broadcast videos
- Video frames taken every 3 seconds over a 5 minute segment

## So how do we classify these?
- Do it in Excel
- Write a website
- Write a program

## Shiny Application
![ShinyApp](img/shiny.png)

# Facial recognition software

## Evaluated software (APIs)

- Animetrics
- Skybiometry
- Microsoft
- Google

## Comparison of features

| Feature        | Google | Microsoft | Skybiometry | Animetrics |
|----------------|--------|-----------|-------------|------------|
| Face features  | ✔      | ✔         | ✔           | ✔          |
| Head angle     | ✔      | ✔         | ✔           | ✔          |
| Basic emotion  | ✔      | ✔         |             |            |
| Face landmarks | ✔      | ✔         |             | ✔          |
| Scene labels   | ✔      |           |             |            |
| 3D points      | ✔      |           |             |            |

## Using the software

> API: Application programming interface

APIs are built to be part of a larger program.

(for example, a camera app)

In order to use them, we need to write scripts to use the API.

## Running the scripts

Detecting faces in each image typically took 3-4 seconds.

To complete all 6406 images, it would take 5-7 hours, per API.

Some APIs had usage limits, and took several days to run.

## Raspberry Pie
![RaspPie](img/raspberry-pie-4.jpg)

## Raspberry Pi
![RPi](img/rasp.png)

## Raspberry Pi
![RaspPi](img/RP.png)

## Several days later...

We have all our data!  
(toward the end of the original project window)

# Preparing the data

## Overlaying face boxes
Check if the software is producing sensible output.
```{r,echo=FALSE, message=FALSE}
ALLmetaIMG <- read.csv("~/github/face-recognition/ALLmetaIMG.csv")

ALLmetaIMG$graphic<-factor(ALLmetaIMG$graphic, levels = 0:1, labels = c("Live image", "Graphic"))
ALLmetaIMG$bg<-factor(ALLmetaIMG$bg, levels = 0:3, labels = c("Crowd", "Court", "Logo wall", "Not applicable"))
ALLmetaIMG$person<-factor(ALLmetaIMG$person, levels = 0:1, labels = c("No Person", "Person"))
ALLmetaIMG$shotangle<-factor(ALLmetaIMG$shotangle, levels = 0:2, labels = c("Player Shoulder Height", "Birds Eye", "Upward Angle"))
ALLmetaIMG$situation<-factor(ALLmetaIMG$situation, levels = 0:5, labels = c("Court in play", "Court player close-up", "Court close-up not player", "Crowd", "Off court close up of player", "Transition"))

ALLmetaIMG$detect<-factor(ALLmetaIMG$detect, levels = 0:3, labels = c("Player", "Other staff on court", "Fan", "None"))
ALLmetaIMG$obscured<-factor(ALLmetaIMG$obscured, levels = 0:1, labels = c("No", "Yes"))
ALLmetaIMG$lighting<-factor(ALLmetaIMG$lighting, levels = 0:2, labels = c("Direct sunlight", "Shaded", "Partially shaded"))
ALLmetaIMG$headangle<-factor(ALLmetaIMG$headangle, levels = 0:3, labels = c("Front on", "Back of head", "Profile", "Other"))
ALLmetaIMG$glasses<-factor(ALLmetaIMG$glasses, levels = 0:1, labels = c("No", "Yes"))
ALLmetaIMG$visorhat<-factor(ALLmetaIMG$visorhat, levels = 0:1, labels = c("No", "Yes"))

library(imager)
library(dplyr)
library(tidyr)

prepareFaceBox <- function(data){
  boxX <- c(data$minX, data$minX, data$maxX, data$maxX, data$minX)
  boxY <- c(data$maxY, data$minY, data$minY, data$maxY, data$maxY)
  data.frame(x = boxX, y = boxY)
}

library(ggplot2)
getManualCount <- function(type, nTotal) {
  return(nTotal[type == "Manual"])
}
ggplotProportion <- function(dataset, factorVar){
  factorVar <- deparse(substitute(factorVar))
  dataset <- dataset %>% filter(matchesManual) %>% group_by_(factorVar, "type") %>% summarise(nTotal=n()) %>% group_by_(factorVar) %>% mutate(ManualCount = getManualCount(type, nTotal)) %>%
    mutate(proportion = nTotal/ManualCount) %>% rename_(xvar = factorVar) %>% filter(type!="Manual")
  ggplot(dataset, aes(x=factor(xvar), y=proportion, group = type, fill=type)) + geom_bar(stat = "identity", position = "dodge") +
    ylab("Proportion of faces matched") + xlab(factorVar)
}

overlayFaceBox <- function(imgList, mergeData, matchBox=TRUE){
  for(i in imgList){
    img <- load.image(paste0("images/", i))
    plot(img)
    faceData <- mergeData %>% filter(file == i)
    if(NROW(faceData) > 0){
      for(face in 1:NROW(faceData)){
        faceBox <- faceData[face,] %>% prepareFaceBox()
        if(matchBox){
          lines(faceBox$x, faceBox$y, col=as.numeric(faceData[face,"boxID"])+1)
        }
        else{
          lines(faceBox$x, faceBox$y, col=as.numeric(faceData[face,"type"])+1)
        }
      }
      if(!matchBox){
        legend("topright", legend = unique(faceData$type), col = unique(as.numeric(faceData$type) + 1), lty = 1)
      }
    }
  }
}
```

## Data is reasonable!

```{r, }
overlayFaceBox("2016_CT6_R01_JMillman_AUS_vs_DSchwartzman_ARG_MS159_clip.0013.png", ALLmetaIMG, FALSE)
```

## Most of the time.

```{r}
overlayFaceBox("2016_HSA_R01_PKohlschreiber_GER_vs_KNishikori_JPN_MS116_clip.0057.png", ALLmetaIMG, FALSE)
```

## Merging the data

Collected data was spread across 6 files, to analyse it, we need to merge it all together.

![Data structure](img/DataLayout.png)

## Matching face boxes

We also need to determine if the boxes from each API match the same face.

This is also important for importing the extra face information from the manual classification.

Comparing every box generated, if the intersection of the box was >10%, they matched.

![box intersection](img/sharedBox.png)

## Matching face boxes
```{r}
overlayFaceBox("2016_HSA_R01_BTomic_AUS_vs_DIstomin_UZB_MS157_clip.0020.png", ALLmetaIMG, TRUE)
```

# Analysing the data

## Shot angles
```{r}
ggplotProportion(ALLmetaIMG, shotangle)
```

## Best shot angle
Player close-up
![closeup](img/closeup.png)

## Worst shot angle
Birds-eye view
![birdseye](img/birdseye.png)

## Hats and visors
As expected, they reduce the detection.
```{r, echo=FALSE}
ggplotProportion(ALLmetaIMG, visorhat)
```

## Classification time
![time](img/recognitionTime.png)

## Face size
![size](img/size.png)

# Conclusions

## The best facial recognition software

Google

- Provided the most accurate face detection
- Provided lots of meta information
- Includes basic emotion recognition
- Well priced

## Room for improvement

Facial recognition software needs work to improve detection of faces, let alone emotion.

Some basic emotion is possible, however the precision is questionable, especially with players expressions during a game.

## Further research

- Use higher definition images
- Obtain and use only best camera angles
- Analyse player microphone audio for emotion
- Use videos for facial recognition

# Q & A