### 2016-12-14 trial.Rmd
Updates

---
title: "Untitled"
author: "Stepahnie Kobakian"
date: "14 December 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(readr)
require(knitr)

require(kfigr)

SceneAttributes<-read_csv("TableCSVs/SceneAttributes.csv")

```

```{r Scene-Attributes-Table, anchor="table"}
kable(SceneAttributes, format = "pandoc", booktabs = TRUE)
```

Table `r figr("Scene-Attributes-Table", TRUE, type="Table")`   Yayyy

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.