### 2017-09-20 ct.tex

%latex.default(ct)%
\begin{table}[!tbp]
\begin{center}
\begin{tabular}{rrrrrrrrrrrrrrrrrrrrlrrrrrlrrrllrrllllllllllllrrrr}
\hline\hline
\multicolumn{1}{c}{tab Animetrics}&\multicolumn{1}{c}{tab Google}&\multicolumn{1}{c}{tab Microsoft}&\multicolumn{1}{c}{tab Skybiometry}&\multicolumn{1}{c}{prop.row Animetrics}&\multicolumn{1}{c}{prop.row Google}&\multicolumn{1}{c}{prop.row Microsoft}&\multicolumn{1}{c}{prop.row Skybiometry}&\multicolumn{1}{c}{prop.col Animetrics}&\multicolumn{1}{c}{prop.col Google}&\multicolumn{1}{c}{prop.col Microsoft}&\multicolumn{1}{c}{prop.col Skybiometry}&\multicolumn{1}{c}{prop.tbl Animetrics}&\multicolumn{1}{c}{prop.tbl Google}&\multicolumn{1}{c}{prop.tbl Microsoft}&\multicolumn{1}{c}{prop.tbl Skybiometry}&\multicolumn{1}{c}{gt}&\multicolumn{1}{c}{rs}&\multicolumn{1}{c}{cs}&\multicolumn{1}{c}{total.n}&\multicolumn{1}{c}{chisq}&\multicolumn{1}{c}{CST}&\multicolumn{1}{c}{chisq.corr}&\multicolumn{1}{c}{fisher.ts}&\multicolumn{1}{c}{fisher.lt}&\multicolumn{1}{c}{fisher.gt}&\multicolumn{1}{c}{print.mcnemar}&\multicolumn{1}{c}{mcnemar}&\multicolumn{1}{c}{mcnemar.corr}&\multicolumn{1}{c}{asr}&\multicolumn{1}{c}{RowData}&\multicolumn{1}{c}{ColData}&\multicolumn{1}{c}{digits}&\multicolumn{1}{c}{max.width}&\multicolumn{1}{c}{vector.x}&\multicolumn{1}{c}{expected}&\multicolumn{1}{c}{prop.chisq}&\multicolumn{1}{c}{resid}&\multicolumn{1}{c}{sresid}&\multicolumn{1}{c}{asresid}&\multicolumn{1}{c}{format}&\multicolumn{1}{c}{cell.layout}&\multicolumn{1}{c}{row.labels}&\multicolumn{1}{c}{percent}&\multicolumn{1}{c}{total.r}&\multicolumn{1}{c}{total.c}&\multicolumn{1}{c}{t Animetrics}&\multicolumn{1}{c}{t Google}&\multicolumn{1}{c}{t Microsoft}&\multicolumn{1}{c}{t Skybiometry}\tabularnewline
\hline
$528$&$ 638$&$505$&$512$&$0.241868987631700$&$0.292258360054970$&$0.231333027943197$&$0.234539624370133$&$0.547717842323652$&$0.325344212136665$&$0.47196261682243$&$0.509452736318408$&$0.1056$&$0.1276$&$0.101$&$0.1024$&$5000$&$2183$&$ 964$&$5000$&FALSE&$$&$$&$$&$$&$$&FALSE&$$&$$&$$&&&$4$&$$&FALSE&FALSE&FALSE&FALSE&FALSE&FALSE&SPSS&FALSE&TRUE&FALSE&TRUE&TRUE&$528$&$ 638$&$505$&$512$\tabularnewline
$436$&$1323$&$565$&$493$&$0.154774582889599$&$0.469648562300319$&$0.200567980120696$&$0.175008874689386$&$0.452282157676349$&$0.674655787863335$&$0.52803738317757$&$0.490547263681592$&$0.0872$&$0.2646$&$0.113$&$0.0986$&$5000$&$2817$&$1961$&$5000$&FALSE&$$&$$&$$&$$&$$&FALSE&$$&$$&$$&&&$4$&$$&FALSE&FALSE&FALSE&FALSE&FALSE&FALSE&SPSS&FALSE&TRUE&FALSE&TRUE&TRUE&$436$&$1323$&$565$&$493$\tabularnewline
\hline
\end{tabular}\end{center}
\end{table}